Robust Winner Determination and Simultaneous Elicitation of Scoring Rules and Preferences

### 2019-02-15 draft.tex
improving MIP optimization

\documentclass[12pt]{article}
\usepackage[top=2cm, bottom=2cm, left=2cm, right=2cm]{geometry}
\geometry{a4paper}
%\usepackage{authblk}

\usepackage{graphicx}
\usepackage{amsmath,amssymb,enumerate,amsthm}
\usepackage{hyperref}
\usepackage{mathrsfs}
\usepackage{algorithm, algpseudocode}
\usepackage{bm}
\usepackage{cleveref}
\usepackage{xcolor}
\usepackage{booktabs}

@@

\algrenewcommand\algorithmicrequire{\textbf{Input:}}
\algrenewcommand\algorithmicensure{\textbf{Output:}}
\newtheorem{claim}{Claim}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}

\title{Robust Winner Determination and Simultaneous Elicitation of Scoring Rules and Preferences}
%\author{}
\author{Olivier Cailloux  \and Beatrice Napolitano \and Paolo Viappiani} %\affil{LIP6\\\email{paolo.viappiani@lip6.fr}
%\author{Olivier Cailloux} 
%\author{Stefano Moretti}%\affil{LAMSADE}
%\author{Beatrice Napolitano}\affil{LAMSADE}

@@

%Recently, the assumption of full a has been questioned by a number of researchers
%In many real situations, however, it may be necessary to reason with partial preferences, as some preferences are not available and too expensive to obtain (with respect to a cognitive or economic cost). 
On the other hand, in several situations it may not be easy to precisely define the voting rule.
In particular, when considering positional scoring rules, it is possible to derive dominance relations (akin to stochastic dominance) that allow to eliminate some alternatives since they will be less preferred than another one for any instantiation of the weights \cite{Stein1994}.Among others, Llamazares and Pe{\~{n}}a \cite{Llamazares2013} have considered  the problem of dealing with underspecified weights in positional scoring rules.
Cailloux and Endriss \cite{Cailloux2014} have proposed elicitation methods for based on partial orders.
Indeed, it is possible that the chair may have some preferences over the desired aggregation method, but may still be not able to formalize the voting rule upfront.
%argued that it might not always be realistic to assume that a voting rule is known precisely.

%We argue that it is possible that the committee wants to precise some preferences about the aggregations.
In this paper we consider that both preferences and the social choice rule can be only partially specified.
We develop methods for computing the minimax-optimal alternative for scoring rules with partial preference information and partial information about the weights.
While previous works have considered either partial preferences or a partially specified aggregation, but we do not know of any work considering both sources of uncertainty at the same time.

@@

We focus on positional scoring rules, that  are a particularly common methods used to aggregate rankings and determine  a winner.
We also address the problem of elicitation, providing incremental elicitation methods to acquire relevant preference information.%The score can be seen as a utility for the society.
%and often captures more information than is needed to determine the winner
We discuss  heuristics that allow us to determine queries that quickly allow to reduce minimax regret.

% the value of certain preference information is often not worth the cost of obtaining it.

%We consider that the committee wants to express preferences for some voting rules
We develop several query strategies that  focus simultaneously on reduction of relevant preference and voting rule uncertainty.
% We must engage in simultaneous elictation of the rule and of the preferences.

@@

%We want to design a questioning strategy to minimize the worst regret in the long run. A questioning strategy indicates, given a partial profile and a set of feasible weights, which question to ask next. A question is either a query to the voter $i$ about her preference relation $\succ_i$ or a query to the committee about the feasible weights.

@@

Possible and necessary winner determination: in typical settings, there will be no necessary winner and too many possible winners.

%\subsection{Related works}
\paragraph{Uncertain scoring rules}
A number of works have dealt with the problem of reasoning with incompletely specified aggregation functions.
%In particular, when considering positional scoring rules, it is possible to derive dominance relations (akin to stochastic dominance) that allow to eliminate some alternatives since they will be less preferred than another one for any instantiation of the weights \cite{Stein1994}.
More recently the characterization of methods for aggregating the uncertainty over the scoring vectors has been studied in \cite{Viappiani2018}.

@@

\section{Background}\label{sec:background}

We introduce the basic concepts.
We assume that there is a set $A$ of $m$ alternatives (for instance products, restaurants, movies, public projects, job candidates, etc.) and $n$ agents (voters); each agent is associated to a preference order.
The goal of the voters is to make a consensus choice, to select a ``winner'' given voter rankings.

%The set of preference orders is called profile and it is denoted by $v$.
The preferences of the agents are supposed to be linear orders (connected, transitive, asymmetric relations) involving the alternatives;
$\pref_i$ denotes the ``real'' preference relation of agent $i$. 
The set $(\pref_1,\ldots,\pref_n)$ is known in the social choice literature as the {\em preference  profile}.
The profile is equivalently represented by $\profile=(v_1,\ldots,v_n)$ where $v_i(j)$ denotes the rank (position) of alternative $j$ in the preference order $\pref_i$. 
With a little abuse of notation, we will use the term profile to refer to either $\profile$ or to the preference relations, depending on the context.
Let $V$ be the set of possible preference profiles (the cartesian product of $n$ linear orders).

A social choice function $f : V \rightarrow 2^A$ associates a profile with a set of winners.
Among the many possible social choice functions, we consider {\em positional scoring rules}, which attach weights to positions; 
%A scoring rule associates each alternative to a score that is given by the sum of points obtained for each voter.
an alternative obtains a score that depends on the rank obtained in each of the preference orders:
\begin{align}
s(x; \profile, \w) = \sum_{j=1}^{n} w_{v_j(x)}
= \sum_{i=1}^{m} \alpha_i w_i . \label{eq:srule}
\end{align}
where $\alpha_i$ is the number of times that alternative $x$ was ranked in the $i$-th position.
The vector $(w_1,\ldots,w_m)$ is called the scoring vector. We assume that the weights constitute a monotonic sequence: $w_{1} \geq w_{2} \geq \ldots \geq w_{m}$.

\medskip
In this work we want to reason about partial preference information.
A partial preference  is encoded by a partial order $\ppref_i$  of voter $i$;
we assume that preference information is truthful, i.e. $a \ppref_i b \implies a \pref_i b$.

A completion of $\ppref_i$ is any linear order $\pref_i$ that extends $\ppref_i$.
Let $C(\ppref_i)$ be the set of completions of $\ppref_i$, that is the set of all complete rankings that extend $\ppref_i$.
An incomplete profile is a set of partial votes %$\textbf{p}=(p_1,\ldots,p_n)$.
$\textbf{p}=(\ppref_1,\ldots,\ppref_n)$.

We let $C(\textbf{p})=C(\ppref_1)\times \ldots \times C(\ppref_n)$ be the set of complete profiles extending $p$.

\medskip
We also assume that the weights of the scoring rule are only partially specified.
We consider a monotonic non-increasing sequence of weights: $w_{1} \geq w_{2} \geq \ldots \geq w_{m}$. Without loss of generality, we assume $w_1=1$ and $w_m=0$
%Without loss of generality, we assume $w_1=1$ and $w_m=0$.
The preferences of the chair are encoded with linear constraints, for instance one may state
that $w_2>0.5$.
We assume the weight sequence to be convex, that means that the difference between the weight of the first position and the weight of the second position is at least as much as the difference between the weights of the second and third positions, etc. 
\[ \forall i \in \{1,\ldots,n\} \;\; w_i - w_{i+1} \geq w_{i+1}-w_{i+2} \iff  w_i - 2 w_{i+1} + w_{i+2} \geq 0 \]
This is a constraint often used when aggregating rankings in sport competitions.
We use $\mathcal{W}$ to denote the set of convex weight vectors.

% Theoretical complexity of partial completition

% We are given  $\ppref_i $,  a partial preference.
% The maximum regret is considered by assuming an adversary can choose the profile
% \begin{align*}
% \PMR(a,b; p) & = \max_{v \in C(p)} s(b; v) - s(a; v) \\
% \MR(a; p) & = \max_{b} \PMR(a,b) 
% \end{align*}
% 
% We then choose the minimax regret optimal alternative:
% \begin{align*}
% \MMR(p) & = \min \PMR(a,b) \\
% a^{*}(p) & \in \arg\min \PMR(a,b) 
% \end{align*}
%Max regret and minimax regret can be computed using independent completion of partial votes for non-decomposable scoring rules.

\section[Minimax regret under partial profile and weight information]{Minimax regret under partial profile and weight information}
\label{sec:mmr}

% This leads to interesting tradeoffs in elicitation
%The regret-minimizing alternative may not be a possible winner for some voting rules.
The weights of a scoring rule can model different preferences of the committee. 
For instance, the weights can control the inclination to favour ``extreme'' alternatives (often at either the top or the bottom of the rankings) at the expenses of ``moderate'' alternatives (that are more consistently in the middle part of the rankings).
However in general it can be difficult to set the weights in an appropriate way.

In this paper, we consider a setting where both the voters' preferences and the preferences of the chair about the voting rule are incomplete.

%\medskip
%The quality of an alternative can be quantified by considering the maximum regret with respect to an adversary that can choose the instantiation of both a complete profile (extending the known preferences of the agents) and of the scoring vectors (associated to the preferences of the committee).

We propose to use the minimax regret as a decision criterion to determine a winner.
Minimax regret \cite{Savage1954} is a decision criterion that has been used for robust optimization under data uncertainty \cite{Kouvelis1997} and as well in decision-making with uncertain utility values  \cite{Salo2001,Boutilier2006}.
Minimax regret \cite{Boutilier2006} has been used for making decisions under utility uncertainty and for preference elicitation.
Lu and Boutilier \cite{Lu2011} have adopted minimax regret for winner determination in social choice with
the preferences of the voters that are only partially known (while the social choice function is known and fixed in advance; it is assumed to be decomposable).

This paper extends the work of  to the simultaneous presence of uncertainty in the agents' preferences and uncertainty in the weights.
Intuitively, the quality of a proposed alternative $a$ is how far from optimal $a$  could be in the worst case, given the current knowledge about the voting rule and about the voters' preferences.

@@

Scoring rules are additively decomposable.
%A scoring rule associates each alternative to a score that is given by the sum of points obtained for each voter.
Let $s(x; v_j,\w)=w_{v_j(x)}$ be the number of points that $x$ obtains in the ranking $v_j$ (see Equation \ref{eq:srule}), that is the weight of position $v_j(x)$.
Obviously we have 
\[ s(x; \profile, \w) = \sum_{j=1}^n s(x; v_j,\w). \]

By exploiting the decomposition of the score in terms of votes, we write the actual regret of choosing $x$ instead of $y$ as:
\[
s(y; \profile,\w) - s(x; \profile, \w) = [\sum_{j=1}^n s(y; v_j,\w) - s(x; v_j,\w)]
\]
and  can rewrite $\PMR$ as follows:
\begin{align*}
\PMR(x,y; \pprofile, W) &= \max_{\w \in W} \max_{\profile \in C(\pprofile)} [ s(y; \profile,\w) - s(x; \profile,\w) ] = \\
&=  \max_{\w \in W} \sum_{j=1}^{n} \max_{v_j \in C(\succ_j^p)} [s(y; v_j,\w) - s(x; v_j,\w)]=\\
&=  \max_{\w \in W} \sum_{j=1}^{n} \max_{v_j \in C(\succ_j^p)} [w_{v_j(y)} - w_{v_j(x)}] \\
\end{align*}
Given our hypothesis, for all pairs of alternatives, there exist completions of the partial preferences $\ppref$ that maximize the PMR and that do not depend on the weights. 
That is, we can exhibit, for each voter $j$, a completion $v_j$ such that $\forall \w, \forall v'_j \in C(\ppref_j): [w_{v_j(y)} - w_{v_j(x)}] \geq [w_{v'_j(y)} - w_{v'_j(x)}]$.

@@

 \item $j \in A^-$: If we know that agent $j$ prefers $x$ to $y$ then the adversary will place as few alternatives as possible between $x$ and $y$.
 We define $U$ as the set of alternatives that are incomparable with $x$ and $y$ according to $\ppref_i$. Since weights are assumed to be convex, the adversary will place all alternatives in $U$ as better than $x$.
\end{itemize}
In summary, some best completion $\pref$ satisfies two properties: (1) for all $c \pinc x$, $c \pref x$; and (2) for all $c \pinc y$, if not $c \pref x \pref y$ then $y \pref c$. These two properties are intuitively understandable by observing that maximizing the PMR requires placing $x$ as low as possible, hence the first property; and $y$ as high as possible, hence the second property, with the exception that sometimes it is not possible to satisfy both desires and noting that the first desire is more important by convexity of the weights. A proof that this reasoning is correct is in \cref{sec:prfCompl}.

@@

Let $\hat{v}$ be the compute profile computed according to the above procedure, and  $\hat{v}_j$ the $j$-th element (the linear order extending the partial order $p_j$ of the $j$-th voter).
The PMR can be written as:
\[ \PMR(x,y) = \sum_{j=1}^n w_{\hat{v}_i(y)} - w_{\hat{v}_i(x)} 
= \sum_{i=1}^m \alpha_i w_i \]
where the coefficient $\alpha_i$ is the number of times $y$ is in the $i-th$ position minus the number of times $x$ is the in the $i$-th position in the rankings of the completed profile $\hat{v}$:
\[ \alpha_{i} = \sum_{j=1}^{n}  I[\hat{v}_{j}(y)=i] - I[\hat{v}_{j}(x)=i]\]
where $I$ is the indicator function.
The linear program is then the following:
\begin{align}
\max & \sum_{i=1}^m \alpha_i w_{i}\\
\text{ s.t. } & w \in W
\end{align}

\subsection{Minimax regret without the convex assumption}

\begin{itemize}	
	\item (will probably go to a later section, or to the appendix)
\end{itemize}

Let $\hat{v}_i$ be the linear order extending $p_i$ according to the above procedure (described in \cite{Lu2011}).
Then PMR can be written as follows:
\[ \PMR(x,y; \pprofile, W) = \max_{\w \in W} \Big \{ \max_{v_j \in C(\succ_j^p)} [w_{v_j(y)} - w_{v_j(x)}]  + \sum_{j \in A^+, j \in A^?} [w_{\hat{v}_j(y)} - w_{\hat{v}_j(x)}] \Big \} \]

We compute pairwise maximum regret by considering binary variables to represent optimization choices related to where to position the alternatives.
We also have numerical variables to represent the weights of the scoring rule.
%Since the objective is to maximize pairwise regret...

 The objective function can be written as:
 \[ \max \sum_{l=1}^n B^j_l w_l -  B^j_l w_l \]

We use enconding trick in order to solve the problem.
We replace the multiplicative terms by new variables.
\[ V_{l}^{j} \sim B^j_l w_l  \]

To impose the desired behavior:
\begin{align*} 
V_{l}^{j} \leq B^j \\
V_{l}^{j} \leq w_l
\end{align*}
The expression $I^j_l w_l$ evaluates to $w_l$ if $I^j_l = 1$, and to $0$ otherwise.

We write $w \in W$ as a shourtcut to represent the constraints that the weights are chosen to be in the feasible set.
%Variable $I^j$ associated with the constraint $0 \leq I \leq |U^j|$.
 Finally, the following is the optimization problem that compute the value $\PMR$.
 The variables are $w$, $V$, $B$,...
\begin{align}
Z := \max & \sum_{l=1}^m  V_{l}^{j,+} - V_{l}^{j,-}\\
\text{ s.t. } & w \in W  \\
& \sum_{B=i_{1}}^{i_{2}} I^{j}_{l} = 1 & \forall j \\
& V_{l}^{j,+} \leq B^j_{l}  & \forall l,j \\
& V_{l}^{j,+} \leq w_l & \forall l,j \\
& V_{l}^{j,-} \geq w + B_{l}^j - 1 & \forall l,j \\
& V_{l}^{j,-} \geq 0 & \forall l,j\\
\end{align} 
The optimization program can be solved by any suitable MILP solver.

@@

If the available information is too limited, the potential error associated with the robust winner (i.e., its max regret) may be unacceptable.

Minimax regret can be used to guide the elicitation process.
As termination condition of elicitation, we can check whether minimax regret is lower than a threshold. If we wish optimality, we can perform elicitation until minimax regret drops to zero.

At each step, the system needs to decide whether to ask a question to one of the agents or ask a question to the chair about the voting rule

In this Section, we describe different strategies to determine informative queries to ask next, with the goal of reducing $\MMR(\pprofile,W)$ quickly.

We propose an regret-based incremental elicitation method...

\paragraph{Query types}
We distinguish between queries asked to the voters and questions asked to the chair.

>>>>

\newif\ifijcai
 \ijcaitrue  %UNCOMMENT this to have the ijcai formatting / COMMENT for draft format

\newif\ifappendix
%\appendixtrue %COMMENT to remove appendix
\ifijcai
 	\input{ijcai_preamble} 
	\author{Olivier Cailloux$^1$	\and Beatrice Napolitano$^1$\and Paolo Viappiani$^{2}$
	\affiliations $^1$ LAMSADE, CNRS and Dauphine University, PSL, Paris, France\\ $^2$ LIP6, UMR 7606, CNRS and Sorbonne Universit\'e, Paris, France
	\emails \{first, second\}@lamsade.fr, paolo.viappiani@lip6.fr
}
\else
	\documentclass[12pt]{article}
	\usepackage[top=2cm, bottom=2cm, left=2cm, right=2cm]{geometry}
	\geometry{a4paper}
	\usepackage{graphicx}
	%\usepackage{authblk}
	\usepackage{booktabs}
	\usepackage{algorithm, algpseudocode}
	\algrenewcommand\algorithmicrequire{\textbf{Input:}}
	\algrenewcommand\algorithmicensure{\textbf{Output:}}
	\author{Various Authors}
\fi

\usepackage{algorithm, algpseudocode}
\usepackage{amsmath,amssymb,enumerate,amsthm}
\usepackage{hyperref}
\usepackage{mathrsfs}

\usepackage{bm}
\usepackage{cleveref}
\usepackage{xcolor}

@@

\newtheorem{claim}{Claim}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}

\title{Robust Winner Determination and Simultaneous Elicitation of Scoring Rules and Preferences}
%\author{Olivier Cailloux  \and Beatrice Napolitano \and Paolo Viappiani}

%\affil{LIP6\\\email{paolo.viappiani@lip6.fr}
%\author{Olivier Cailloux} 
%\author{Stefano Moretti}%\affil{LAMSADE}
%\author{Beatrice Napolitano}\affil{LAMSADE}

@@

%Recently, the assumption of full a has been questioned by a number of researchers
%In many real situations, however, it may be necessary to reason with partial preferences, as some preferences are not available and too expensive to obtain (with respect to a cognitive or economic cost). 
On the other hand, in several situations it may not be easy to precisely define the voting rule.
In particular, when considering positional scoring rules, it is possible to derive dominance relations (akin to stochastic dominance) that allow to eliminate some alternatives since they will be less preferred than another one for any instantiation of the weights \cite{Stein1994}.
Among others, Llamazares and Pe{\~{n}}a \cite{Llamazares2013} have considered  the problem of dealing with underspecified weights in positional scoring rules.
Cailloux and Endriss \cite{Cailloux2014} have proposed elicitation methods for based on partial orders.
Indeed, it is possible that the chair may have some preferences over the desired aggregation method, but may still be not able to formalize the voting rule upfront.
%argued that it might not always be realistic to assume that a voting rule is known precisely.

%We argue that it is possible that the committee wants to precise some preferences about the aggregations.
In this paper we consider that both preferences and the social choice rule can be only partially specified.
We develop methods for computing the minimax-optimal alternative for scoring rules with partial preference information and partial information about the weights.
While previous works have considered either partial preferences or a partially specified aggregation, we do not know of any work considering both sources of uncertainty at the same time.

@@

We focus on positional scoring rules, that  are a particularly common methods used to aggregate rankings and determine  a winner.
We also address the problem of elicitation, providing incremental elicitation methods to acquire relevant preference information. 
%The score can be seen as a utility for the society.
%and often captures more information than is needed to determine the winner
We discuss  heuristics that allow us to determine queries that quickly allow to reduce minimax regret.
We develop several query strategies that  focus simultaneously on reduction of relevant preference and voting rule uncertainty.
% the value of certain preference information is often not worth the cost of obtaining it.

%We consider that the committee wants to express preferences for some voting rules

% We must engage in simultaneous elictation of the rule and of the preferences.

@@

%We want to design a questioning strategy to minimize the worst regret in the long run
% A questioning strategy indicates, given a partial profile and a set of feasible weights, which question to ask next. A question is either a query to the voter $i$ about her preference relation $\succ_i$ or a query to the committee about the feasible weights.

@@

% we define a model that allows simultaneous elicitation of the voting rule and user preferences, making appropriate tradeoffs be- tween the two types of information

%\subsection{Related works}
\paragraph{Text to be placed somewhere:}
%A number of works have dealt with the problem of reasoning with incompletely specified aggregation functions.
%In particular, when considering positional scoring rules, it is possible to derive dominance relations (akin to stochastic dominance) that allow to eliminate some alternatives since they will be less preferred than another one for any instantiation of the weights \cite{Stein1994}.
More recently the characterization of methods for aggregating the uncertainty over the scoring vectors has been studied in \cite{Viappiani2018}.

\medskip
Using {\em maximum regret} to quantify the worst-case error, the alternative that minimize this error is selected as winner, providing us with a form of robust optimization.

\medskip
Possible and necessary winner determination \cite{Xia2008}: in typical settings, there will be no necessary winner and too many possible winners.

@@

\section{Social choice with partial information}\label{sec:background}

In this Section we introduce some basic concepts.
We assume that there is a set $A$ of $m$ alternatives (for instance products, restaurants, movies, public projects, job candidates, etc.) and $n$ agents (voters); each agent is associated to a preference order.
The goal of the voters is to make a consensus choice, in order to select a ``winner'' given voter rankings.

%The set of preference orders is called profile and it is denoted by $v$.
The preferences of the agents are supposed to be linear orders (connected, transitive, asymmetric relations) involving the alternatives;
$\pref_i$ denotes the ``real'' preference relation of agent $i$. 
The set $(\pref_1,\ldots,\pref_n)$ is known in the social choice literature as the {\em preference  profile}.
The profile is equivalently represented by $\profile=(v_1,\ldots,v_n)$ where $v_i(j)$ denotes the rank (position) of alternative $j$ in the preference order $\pref_i$. 
With a little abuse of notation, we will use the term profile to refer to either $\profile$ or to the preference relations, depending on the context.

Let $V$ be the set of possible preference profiles (the cartesian product of $n$ linear orders).
A social choice function $f : V \rightarrow 2^A$ associates a profile with a set of winners.
Among the many possible social choice functions, we consider {\em positional scoring rules}, which attach weights to positions; 
%A scoring rule associates each alternative to a score that is given by the sum of points obtained for each voter.
an alternative obtains a score that depends on the rank obtained in each of the preference orders:
\begin{align}
s(x; \profile, \w) = \sum_{j=1}^{n} w_{v_j(x)}
= \sum_{i=1}^{m} \alpha^{x}_i w_i . \label{eq:srule}
\end{align}
where $\alpha^{x}_i$ is the number of times that alternative $x$ was ranked in the $i$-th position.
The winner is the alternative with highest score.
The vector $(w_1,\ldots,w_m)$ is called the scoring vector.
%We assume that the weights constitute a monotonic sequence: $w_{1} \geq w_{2} \geq \ldots \geq w_{m}$.
In this work we want to reason about partial preference information.
A partial preference  is encoded by a partial order $\ppref_i$  of voter $i$; in this work we assume that preference information is truthful, i.e. $a \ppref_i b \implies a \pref_i b$.
An incomplete profile is a set of partial votes  $\textbf{p}=(\ppref_1,\ldots,\ppref_n)$.

A completion of $\ppref_i$ is any linear order $\pref_i$ that extends $\ppref_i$.
Let $C(\ppref_i)$ be the set of completions of $\ppref_i$, that is the set of all complete rankings that extend $\ppref_i$.
We let $C(\textbf{p})=C(\ppref_1)\times \ldots \times C(\ppref_n)$ be the set of complete profiles extending $p$.

\medskip
We also assume that the weights of the scoring rule are only partially specified.
We consider a monotonic non-increasing sequence of weights: $w_{1} \geq w_{2} \geq \ldots \geq w_{m}$; this is a natural assumption, as it is better to be ranked first than second, second than third, etc. 
Without loss of generality, we assume $w_1=1$ and $w_m=0$.

The weights of a scoring rule can model different preferences of the committee. 
For instance, the weights can control the inclination to favour ``extreme'' alternatives (often at either the top or the bottom of the rankings) at the expenses of ``moderate'' alternatives (that are more consistently in the middle part of the rankings).
However, we argue that in general it can be difficult to set the weights in an appropriate way.

An important class of scoring rule is the one composed of weights that are a convex sequence, meaning that the difference between the weight of the first position and the weight of the second position is at least as much as the difference between the weights of the second and third positions, etc.
\begin{align} \forall i \in \{1,\ldots,n-2\} \;\; & w_i - w_{i+1} \geq w_{i+1}-w_{i+2} \iff  \\
& w_i - 2 w_{i+1} + w_{i+2} \geq 0 \label{eq:convexity}
\end{align}
The constraint above is often used when aggregating rankings in sport competitions.
We use $\mathcal{W}$ to denote the set of convex weight vectors.

Additionally the chair (the person or the organization that is supervising the voting process) may be able to specify additional preferences about how the social choice function should behave.
In this work we assume that preferences of the chair are encoded with linear constraints on the vector $\w$, relating the value of the weights of different positions.
Of course it may be difficult for real decision makers to state preferences about the voting rule in such an abstract way, unless perhaps in some limited cases (a specific instance of these preferences may consist in giving upper or lower bound on the value of $w_{i}$ for position $i$).
These additional preferences can be elicited by asking questions about concrete profiles, for instance, by showing a complete profile of a small synthetic election and asking who should be elected in this case.
We discuss queries asked to the chairs and strategies to select queries later in Section \ref{sec:elicit}.

% Theoretical complexity of partial completition
% We are given  $\ppref_i $,  a partial preference.
% The maximum regret is considered by assuming an adversary can choose the profile
% \begin{align*}
% \PMR(a,b; p) & = \max_{v \in C(p)} s(b; v) - s(a; v) \\
% \MR(a; p) & = \max_{b} \PMR(a,b) 
% \end{align*}
% 
% We then choose the minimax regret optimal alternative:
% \begin{align*}
% \MMR(p) & = \min \PMR(a,b) \\
% a^{*}(p) & \in \arg\min \PMR(a,b) 
% \end{align*}
%Max regret and minimax regret can be computed using independent completion of partial votes for non-decomposable scoring rules.

\section[Minimax regret under partial profile and weight information]{
Robust winner determination}
\label{sec:mmr}

% This leads to interesting tradeoffs in elicitation
%The regret-minimizing alternative may not be a possible winner for some voting rules.
In this paper, we consider a setting where both the voters' preferences and the preferences of the chair about the voting rule are incomplete.
%\medskip
%The quality of an alternative can be quantified by considering the maximum regret with respect to an adversary that can choose the instantiation of both a complete profile (extending the known preferences of the agents) and of the scoring vectors (associated to the preferences of the committee).

As a decision criterion to determine a winner, we propose to use the minimax regret. 
Minimax regret \cite{Savage1954} is a decision criterion that has been used for robust optimization under data uncertainty \cite{Kouvelis1997} and as well in decision-making with uncertain utility values  \cite{Salo2001,Boutilier2006}.
%Minimax regret \cite{Boutilier2006} has been used for making decisions under utility uncertainty and for preference elicitation.
Lu and Boutilier \cite{Lu2011} have adopted minimax regret for winner determination in social choice with
the preferences of the voters that are only partially known, while the social choice function is known and fixed in advance.

This paper tackles the more challenging setting of the simultaneous presence of uncertainty in the agents' preferences and uncertainty in the weights.
Intuitively, the quality of a proposed alternative $a$ is how far from optimal $a$  could be in the worst case, given the current knowledge about the voting rule and about the voters' preferences.

@@

Scoring rules are additively decomposable.
%A scoring rule associates each alternative to a score that is given by the sum of points obtained for each voter.
Let $s(x; v_j,\w)=w_{v_j(x)}$ be the number of points that $x$ obtains in the ranking $v_j$ (see Equation \ref{eq:srule}), that is given by weight of position $v_j(x)$.
Obviously we have 
\[ s(x; \profile, \w) = \sum_{j=1}^n s(x; v_j,\w). \]

By exploiting the decomposition of the score in terms of votes, we write the actual regret of choosing $x$ instead of $y$ as:
\[
s(y; \profile,\w) - s(x; \profile, \w) = [\sum_{j=1}^n s(y; v_j,\w) - s(x; v_j,\w)]
\]
and  can rewrite $\PMR$ as follows:
\begin{align*}
& \PMR(x,y; \pprofile, W) = \max_{\w \in W} \max_{\profile \in C(\pprofile)} [ s(y; \profile,\w) - s(x; \profile,\w) ] = \\
& =  \max_{\w \in W} \sum_{j=1}^{n} \max_{v_j \in C(\succ_j^p)} [s(y; v_j,\w) - s(x; v_j,\w)]=\\
& =  \max_{\w \in W} \sum_{j=1}^{n} \max_{v_j \in C(\succ_j^p)} [w_{v_j(y)} - w_{v_j(x)}] \\
\end{align*}
Given our hypothesis, for all pairs of alternatives, there exist completions of the partial preferences $\ppref$ that maximize the PMR and that do not depend on the weights. 
That is, we can exhibit, for each voter $j$, a completion $v_j$ such that $\forall \w, \forall v'_j \in C(\ppref_j): [w_{v_j(y)} - w_{v_j(x)}] \geq [w_{v'_j(y)} - w_{v'_j(x)}]$.

@@

 \item $j \in A^-$: If we know that agent $j$ prefers $x$ to $y$ then the adversary will place as few alternatives as possible between $x$ and $y$.
Define $U$ as the set of alternatives that are incomparable with $x$ and $y$ according to $\ppref_i$. Since weights are assumed to be convex, the adversary will place all alternatives in $U$ as better than $x$.
\end{itemize}
In summary, some best completion $\pref$ satisfies two properties: (1) for all $c \pinc x$, $c \pref x$; and (2) for all $c \pinc y$, if not $c \pref x \pref y$ then $y \pref c$. These two properties are intuitively understandable by observing that maximizing the PMR requires placing $x$ as low as possible, hence the first property; and $y$ as high as possible, hence the second property, with the exception that sometimes it is not possible to satisfy both goals and noting that the first goal is more important by convexity of the weights. 
A proof that this reasoning is correct is in \cref{sec:prfCompl}.

@@

Let $\hat{v}$ be the compute profile computed according to the above procedure, and  $\hat{v}_j$ the $j$-th element (the linear order extending the partial order $p_j$ of the $j$-th voter).
The PMR can be written as:
\[ \PMR(x,y; \pprofile, W) = \sum_{j=1}^n w_{\hat{v}_i(y)} - w_{\hat{v}_i(x)} 
= \sum_{i=1}^m (\alpha_{i}^{y} - \alpha_{i}^{x}) w_i \]
where the coefficient $ (\alpha_{i}^{y} - \alpha_{i}^{x})$ is the number of times $y$ is in position $i$ minus the number of times $x$ is the in position $i$ in the rankings of the completed profile $\hat{v}$:
$\alpha_{i}^{x} = \sum_{j=1}^{n}  I[\hat{v}_{j}(x)=i]$.
%\[ \alpha_{i} = \sum_{j=1}^{n}  I[\hat{v}_{j}(y)=i] - I[\hat{v}_{j}(x)=i]\]
where $I$ is the indicator function.
Then, what we have to to compute pairwise max regret is to compute the solution of the following linear program :
\begin{align*}
\PMR(x,y; \pprofile,W) = \max_{\w} & \sum_{i=1}^m \alpha_i w_{i}\\
\text{ s.t. } & \w \in W \\
& \text{Equation } (\ref{eq:convexity})
\end{align*}
The result of the optimization is the value of the pairwise max regret $\PMR$.

\subsection{Minimax regret without the convex assumption}
{\em (will probably go to a later section, or to the appendix) }

\medskip
Let $\hat{v}_i$ be the linear order extending $p_i$ according to the above procedure (described in \cite{Lu2011}).
Then PMR can be written as follows:
\begin{align*}
 &\PMR(x,y; \pprofile, W) =\\ 
 &\max_{\w \in W} \Big \{ \max_{v_j \in C(\succ_j^p)} [w_{v_j(y)} \!-\! w_{v_j(x)}]  \!+\!	 \sum_{j \in A^+ \cup  A^?} [w_{\hat{v}_j(y)} \!-\! w_{\hat{v}_j(x)}] \Big \} 
 \end{align*}

%We compute pairwise maximum regret by considering binary variables $\{ B_{i} \}_{i=1,\ldots,m}$ to represent optimization choices related to where to position the alternatives.
We introduce two set of binary variables $B^{+}_{i,j}$ and $B^{-}_{i,j}$  for each voter $i$ and each position $j$.
The variable $B_{i,j}^{+}$ encodes the fact that the alternative $y$ is placed in position $j$ in the ranking of agent $i$; while  $B_{i,j}^{-}$ encodes the position of the alternative $x$.

We also have numerical variables to represent the weights of the scoring rule.
Since each alternative needs to be placed exactly in one place for each agent, we adopt the constraints
$\sum_{i=1}^{m} B_{i,j}^{+} = 1$ and $\sum_{i=1}^{m} B_{i,j}^{-} = 1$.
%Since the objective is to maximize pairwise regret...

The score of alternative $y$ can be written as $\sum_{i=1}^{n} \sum_{j=1}^{m} w_{j} B_{i,j}^{+}$.
Our objective function is now:
 \[ \max \sum_{i=1}^{n} \sum_{j=1}^{m} B_{i,j}^{+} w_j- B_{i,j}^{-}  w_j \]

We use integer programming enconding tricks in order to linearize the problem formulation.
We introduce yet another set of variables  $V_{i,j}^{+} $  and $V_{i,j}^{-}$ % the multiplicative terms by new variables.
and we introduce constraints so that $V_{i,j}^{+} = B^{+}_{i,j} w_j$.
To impose the desired behavior, we set $V_{l}^{j} \leq B^j $ and $V_{l}^{j} \leq w_l$.
%\begin{align*} 
%V_{l}^{j} \leq B^j \\
%V_{l}^{j} \leq w_l
%\end{align*}
%The expression $B^j_l w_l$ evaluates to $w_l$ if $I^j_l = 1$, and to $0$ otherwise.

%Variable $I^j$ associated with the constraint $0 \leq I \leq |U^j|$.
 Finally, the following is the optimization problem that compute the value $\PMR$.
 %The decision variables are $w$, $V$, $B$,...

We therefore obtain the following mixed integer linear program:
\begin{align}
\max & \sum_{i=1}^n  \sum_{j=1}^m  V_{i,j}^{+} - V_{i,j}^{-}\\
\text{ s.t. } & w \in W  \\
& \sum_{i=t_{1}}^{t_{2}} B_{i,j}^{+} = 1 & \forall j \\
& V_{i,j}^{+} \leq B_{i,j}^{+}  & \forall i,j \\
& V_{i,j}^{+} \leq w_j & \forall i,j \\
& V_{i,j}^{-} \geq w + B_{i,j}^{-} - 1 & \forall i,j \\
& V_{i,j}^{-} \geq 0 & \forall i,j
\end{align} 
%We write $w \in W$ as a shourtcut to represent the constraints that the weights are chosen to be in the feasible set.

The optimization program can be solved by any suitable MILP solver.

@@

We propose an incremental elicitation method based on minimax regret.
At each step, the system may either ask a question to one of the agents, or ask a question to the chair about the voting rule. 
The goal is to acquire relevant information to reduce minimax regret as quick as possible.
%Minimax regret can be used to guide the elicitation process.
As termination condition of elicitation, we can check whether minimax regret is lower than a threshold; if we wish optimality, we can perform elicitation until minimax regret drops to zero.

The remainder of this Section is structured as follows.
First of all, we discuss the different types of queries that can be asked to the voters and to the chair, and the way responses are handled.
Then, we describe different strategies to determine informative queries to ask next, with the goal of reducing $\MMR(\pprofile,W)$ quickly.
